### 2024-07-24 tex/cover_latter.tex
after resubmission

\documentclass{letter}
\usepackage{hyperref}
\usepackage{fullpage}

\signature{Dr Geraint Palmer}
\begin{document}

\begin{letter}{}
\opening{Dear editors,}

My name is Dr Geraint Palmer, lecturer at the School of Mathematics at Cardiff University. I am writing on behalf on of a team of authors as we wish to submit a research article entitled \textit{``Optimising Heterogeneous Ambulance Fleet Allocations in Jakarta''} for consideration in the European Journal of Operational Research.

In this paper, we propose a modelling approach to consider where to optimally locate different types of emergency response vehicles in order to maximise patient outcomes within a heterogeneous population. We use this modelling approach within a heuristic algorithm to optimally allocate the vehicles, and use a novel discrete event simulation to evaluate the found allocations.
This paper is a continuation and extension of previously published work, namely \href{https://www.sciencedirect.com/science/article/pii/S0305048312000436?casa_token=1SiRonMwOYIAAAAA:AnZOLnmzNdYATlXDMYjqnQHEZX701o1eaBypnyMpgQbqoR62Z57aX1SXAadw4GJAku8gNvc}{\textit{``Ambulance allocation for maximal survival with heterogeneous outcome measures''}} by Knight, Harper and Smith.
This paper extends that work of that paper in four ways:

\begin{itemize}
  \item we extend the model to include heterogeneous fleets, that is both primary transport vehicles and secondary response vehicles;
  \item we overcome one problem identified in the first paper, that is the circular nature of the vehicle utilisations, by numerically solving implicit equations relating allocation and utilisation;
  \item we develop a sequential discrete event simulation of the calls and emergency response vehicles' logic to evaluate the found allocations on a range of KPIs;
  \item we apply the work to the real life situation in the city of Jakarta, in a developing country.
\end{itemize}

Thank you for your consideration of this article.

\closing{Yours Faithfully,}

\end{letter}
\end{document}